---
title: "Homework 4"
author: "Lizzy Gibson"
date: "October 24, 2017"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(janitor)
```

## Read and Merge 538 Data
```{r}
pols <- read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  clean_names() %>% 
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  mutate(date = lubridate::floor_date(mon,"month")) %>% 
  select(-mon) %>% 
  arrange(date)

gdp <- read_csv("./fivethirtyeight_datasets/GDP.csv", na = ".") %>% 
  clean_names() %>% 
  select(date, gdp_value = value) %>% 
  mutate(date = lubridate::floor_date(date,"month"))

five38 <- inner_join(pols, gdp, by = "date") %>% 
  select(date, everything())
```

### View Data
```{r view}
five38
```

## GDP Value over Time

```{r_five38_plot}
five38 %>% 
   mutate(president = forcats::fct_recode(president,
                                         "GOP" = "gop",
                                         "Dem" = "dem")) %>%
  ggplot(aes(x = date, y = gdp_value, color = president)) + 
  geom_col(aes(fill = president)) +
  labs(x = "Date", y = "GDP Value", title = "Figure 2: GDP Value Over Time for Each Presidential Party")
```

From Figure 2 above, it can be seen that whether the president was a democrat or republican, the GDP was constantly fluctuating. The highest GDP was attained in the early 1950's when a democrat, President Truman, was in office. The next highest GDP was attained when yet another democrat, President Carter, was in office in the late 1970's. The largest fluctuation occurred in the late 1950's, during (republican) President Eisenhower's term. There appears to be an overall decrease in GDP from the 1980's to 2014, where it reached an all-time low with (democratic) President Obama's term in office. The nation's financial stability, which relates heavily to politics and current events such as wars, is constantly changing even within a single president's term.  


